---
title: "processing"
author: "Indra Boving"
date: "2024-04-29"
output: html_document
---

TO DO: 

- Entered wrong data for 4/29 post (same as pre, check that)
- temp and rh data from doing gmin, imageJ on the nosmoke leaves (do they exist?)
- figure out which trees we should have for ACi curves from 5/1 - 5/3; seems like we're missing some smoke ones? Should be 6 and 6? 

```{r}
library(tidyverse)
install.packages("readxl")
library(readxl)

install.packages(c("units", "stringi", "jsonify", "xml2", "tidyxl", "tibble", "vctrs", "devtools"))
devtools::install_local("gasanalyzer-master.zip")

#custon ggplot: 
ggplot <- function(...) { ggplot2::ggplot(...) + 
    theme(panel.background = element_rect(fill='white', colour='black'), # Make background white and border black
          panel.grid.major = element_blank(),  # Hide major gridlines
          panel.grid.minor = element_blank(),# Hide minor gridlines
          strip.background = element_rect(fill = "white"))  # Make background white
}

theme_minimal <- theme_minimal() + 
  theme(panel.background = element_rect(fill='white', colour='black'), # Make background white and border black
        panel.grid.major = element_blank(),  # Hide major gridlines
        panel.grid.minor = element_blank(),# Hide minor gridlines
        strip.background = element_rect(fill = "white")) 

```

Using code from this: 
```{r setup, include=FALSE}
# knitr::opts_chunk$set(echo = TRUE)
# # install.packages("remotes")
# remotes::install_github("muir-lab/licorer")
# install.packages("licorer")
# 
# data <- read_li6800_excel(here::here("raw-data","2024-04-25-1052_logdata_smoke_test_r1"), decimal = ".")
```

#-----
#LICOR DATA: 

- First make sure all cells in the excel sheet are values, not equations (select all, copy, then paste special as values)
```{r}
# Load required libraries
library(readxl)
library(dplyr)
library(lubridate)

# Define the folder path containing Excel files
folder_path <- here::here("raw-data","licor tests", "2024-03-19 segi smoke", "licor_forR")


# List all Excel files in the folder
files <- list.files(path = folder_path, pattern = "*.xlsx", full.names = TRUE)

# Initialize an empty list to store individual dataframes
df_list <- list()

# Loop through each file in the folder
for (file in files) {
  
  # Read in the Excel sheet, skipping formulas and reading only values
  data <- read_excel(file, sheet = "Measurements", col_types = "text")
  
  # Skip the first 13 rows manually
  data <- data[-c(1:12), ]
  
  # Combine rows 1, 2, and 3 (corresponding to rows 14, 15, and 16 in original) to create column names
  combined_row <- paste(data[1, ], data[2, ], sep = "_")
  
  # Remove the rows 1, 2, and 3 after combining them (they are now column headers)
  data <- data[-c(1, 2), ]
  
  # Assign combined row as the new column names
  colnames(data) <- combined_row
  
  # Remove instances of _NA from column names
  colnames(data) <- gsub("_NA", "", colnames(data))
  
  # Add the processed dataframe to the list
  df_list[[file]] <- data
}
library(hms)

# Combine all dataframes into one large dataframe
final_data <- bind_rows(df_list) %>% 
  mutate(date_time= as.POSIXct(date, format = "%Y%m%d %H:%M:%S"),
         date_time = lubridate::ymd_hms(date), 
         date = date(date_time), 
         time = as_hms(date_time)) %>% 
  janitor::clean_names() %>%
  mutate(across(
    .cols = where(is.character) & !all_of(c("plant_id", "na_cat")),  # Exclude 'plant_id' and 'NA_cat'
    .fns = as.numeric), 
    type = na_cat) %>% 
  drop_na(date) %>% 
  filter(!(date %in% c("2024-04-22"))) %>% 
  filter(a_mmol_m_2_s_1 < 50 ) %>% #remove one outlier 
  filter(a_mmol_m_2_s_1 > 0) %>% #remove one outlier 
  mutate(spot_aci = case_when(
    type %in% c("pre", "post", "post 4 days", "post 9 days") ~ "spot measurements", 
    TRUE ~ "ACi curves"
  )) %>% 
  group_by(date, type, plant_id, spot_aci) %>% 
  mutate(A = mean(a_mmol_m_2_s_1, na.rm = T), 
         gsw = mean(gsw_mol_m_2_s_1, na.rm = T)) %>% 
  ungroup() %>%
  mutate(plant_id = case_when(
    # Replace 'sm_' with 'smoke_' and ensure the number is always after 'smoke_'
    str_detect(plant_id, "^sm_") ~ str_replace(plant_id, "sm_", "smoke_"),
    str_detect(plant_id, "^\\d+_smoke") ~ str_replace(plant_id, "(\\d+)_smoke", "smoke_\\1"),
    str_detect(plant_id, "_smoke\\d+") ~ str_replace(plant_id, "_smoke(\\d+)", "smoke_\\1"),
    
    # Replace 'no_' with 'nosmoke_' and ensure the number is always after 'nosmoke_'
    str_detect(plant_id, "^no_") ~ str_replace(plant_id, "no_", "nosmoke_"),
    str_detect(plant_id, "^\\d+_nosmoke") ~ str_replace(plant_id, "(\\d+)_nosmoke", "nosmoke_\\1"),
    str_detect(plant_id, "_nosmoke\\d+") ~ str_replace(plant_id, "_nosmoke(\\d+)", "nosmoke_\\1"),
    
    # Case where the number is already after 'smoke_' or 'nosmoke_' (no changes)
    str_detect(plant_id, "smoke_\\d+") ~ plant_id,
    str_detect(plant_id, "nosmoke_\\d+") ~ plant_id,
    
    # Leave everything else untouched
    TRUE ~ plant_id
  ))%>%
  mutate(smoke_status = case_when(
    str_detect(plant_id, "^smoke_") ~ "smoke",
    str_detect(plant_id, "^nosmoke_") ~ "nosmoke",
    TRUE ~ NA_character_  # Set as NA for any other cases
  ))

unique(final_data$plant_id)
```

```{r}
day1 <- final_data %>% 
  filter(date %in% c("2024-04-25"))

unique(day1$plant_id)
```
#####Assimilation pre and post: 

```{r}
level_order <- c('pre', 'post')

final_data %>% 
  select(date, A, spot_aci, type, plant_id, smoke_status) %>% 
  distinct() %>% 
  filter(date %in% c("2024-04-29", "2024-04-30", "2024-04-26", 
                     "2024-04-25")) %>% 
  filter(smoke_status == "smoke") %>% 
  filter(spot_aci == "spot measurements") %>% 
  ggplot(aes(y = A, x = factor(type, level = level_order), color = type)) +
  geom_jitter(alpha = .3) +
  stat_summary(
    fun = mean,                         # Plot the mean as the dot
    fun.data = mean_sdl,                 # Plot error bars using standard error
    geom = "pointrange",                # Point with error bars
    size = 1,                          # Size of the dot
    position = position_dodge(width = 1)
  )  +
  theme(legend.position = "none") +
  scale_color_manual(values = c( "lightpink", "darkgreen")) +
  facet_wrap(~date ,nrow = 1) +
  labs(y = "Assimilation µmol m⁻² s⁻¹", 
       x = "Time")
  
```
#####Assimilation control vs. smoke: 

```{r, fig.height=3, fig.width=8}
level_order <- c('smoke', 'nosmoke')

df <- final_data %>% 
  select(date, gsw, A, spot_aci, type, plant_id, smoke_status) %>% 
   distinct() %>% 
  filter(!(date %in% c("2024-05-01", "2024-05-02", "2024-05-03", "2022-03-03")))
  #                    "2024-04-25", "2024-05-13", "2024-05-08", "2024-05-07")) #%>% 
  #filter(date %in% c("2024-04-29", "2024-04-30", "2024-04-26", 
  #                    "2024-04-25", "2024-05-13", "2024-05-08", "2024-05-07")) 

df %>% 
  #filter(smoke_status == "smoke") %>% 
  filter(!(type %in% c("post"))) %>% 
  #filter(spot_aci == "spot measurements") %>% 
  ggplot(aes(y = gsw, color = factor(smoke_status, level = level_order), x = date)) +
  geom_jitter(alpha = .3) +
  stat_summary(
    fun = mean,                         # Plot the mean as the dot
    fun.data = mean_sdl,                 # Plot error bars using standard error
    geom = "pointrange",                # Point with error bars
    size = .75,                          # Size of the dot
    position = position_dodge(width = .5)
  ) +
  #facet_wrap(~date ,nrow = 1) +
  labs(y = "Assimilation (µmol m⁻² s⁻¹)", 
       x = "Time", 
       color = "Treatment")
  
```

#####GSW pre and post: 

```{r}
level_order <- c('pre', 'post')

final_data %>% 
  select(date, gsw, spot_aci, type, plant_id, smoke_status) %>% 
  distinct() %>% 
  filter(date %in% c("2024-04-29", "2024-04-30", "2024-04-26", 
                     "2024-04-25")) %>% 
  filter(smoke_status == "smoke") %>% 
  filter(spot_aci == "spot measurements") %>% 
  ggplot(aes(y = gsw, x = factor(type, level = level_order), color = type)) +
  geom_jitter(alpha = .3) +
  stat_summary(
    fun = mean,                         # Plot the mean as the dot
    fun.data = mean_sdl,                 # Plot error bars using standard error
    geom = "pointrange",                # Point with error bars
    size = 1,                          # Size of the dot
    position = position_dodge(width = 1)
  )  +
  theme(legend.position = "none") +
  scale_color_manual(values = c( "lightpink", "darkgreen")) +
  facet_wrap(~date ,nrow = 1) +
  labs(y = "Stomatal Conductance (mol m⁻² s⁻¹)", 
       x = "Time") 
  
```

#####GSW control vs. smoke: 

```{r, fig.height=3, fig.width=8}
level_order <- c('nosmoke', 'smoke')

df <- final_data %>% 
  select(date, gsw, A, spot_aci, type, plant_id, smoke_status) %>% 
   distinct() %>% 
  filter(!(date %in% c("2024-05-01", "2024-05-02", "2024-05-03", "2022-03-03")))
  #                    "2024-04-25", "2024-05-13", "2024-05-08", "2024-05-07")) #%>% 
  #filter(date %in% c("2024-04-29", "2024-04-30", "2024-04-26", 
  #                    "2024-04-25", "2024-05-13", "2024-05-08", "2024-05-07")) 

df %>% 
  #filter(smoke_status == "smoke") %>% 
  filter(!(type %in% c("post"))) %>% 
  #filter(spot_aci == "spot measurements") %>% 
  ggplot(aes(y = gsw, x = factor(smoke_status, level = level_order), color = smoke_status)) +
  geom_jitter(alpha = .3) +
  stat_summary(
    fun = mean,                         # Plot the mean as the dot
    fun.data = mean_sdl,                 # Plot error bars using standard error
    geom = "pointrange",                # Point with error bars
    size = 1,                          # Size of the dot
    position = position_dodge(width = 1)
  ) +
  facet_wrap(~date ,nrow = 1) +
  labs(y = "Stomatal Conductance (mol m⁻² s⁻¹)", 
       x = "Time")

head(df)
unique(final_data$date)

df %>% 
  #filter(smoke_status == "smoke") %>% 
  filter(!(type %in% c("post"))) %>% 
  #filter(spot_aci == "spot measurements") %>% 
  ggplot(aes(y = gsw, color = factor(smoke_status, level = level_order), x = date)) +
  geom_jitter(alpha = .3) +
  stat_summary(
    fun = mean,                         # Plot the mean as the dot
    fun.data = mean_sdl,                 # Plot error bars using standard error
    geom = "pointrange",                # Point with error bars
    size = .75,                          # Size of the dot
    position = position_dodge(width = .5)
  ) +
  #facet_wrap(~date ,nrow = 1) +
  labs(y = "Stomatal Conductance (mol m⁻² s⁻¹)", 
       x = "Time", 
       color = "Treatment")

```

#------

#####VCmax
Aci curves (the ones that are good):

```{r}
aci_post <- final_data %>% 
  filter(#spot_aci %in% c("ACi Curves") &
           date %in% c("2024-05-03", 
                       "2024-05-02", 
                       "2024-05-04")) %>% 
  drop_na(plant_id) %>% 
  filter(ci_mmol_mol_1 > 200) %>% 
  mutate(day = 1, 
         licor = "anderegg", 
         id = plant_id, 
         Photo = a_mmol_m_2_s_1, 
         Ci = ci_mmol_mol_1,
         Tleaf = tleaf_c, 
         PARi = q_mmol_m_2_s_1) %>% 
  select(day, licor, id, Photo, Ci, Tleaf, PARi, elapsed_s) %>% 
  filter(!(id %in% c("smoke_7", "smoke_10"))) %>% 
  filter(!(id %in% c("smoke_12") & elapsed_s <= 6500)) %>% 
  filter(!(id %in% c("nosmoke_4") & elapsed_s <= 10200))%>% 
  filter(Ci < 2500) 


##I think smoke_12 is two different trees that are labeled wrong... 
aci_post %>% 
 # filter(id == "nosmoke_4") %>% 
  #filter(Ci < 2000) %>% 
  ggplot(aes(y = Photo, x = Ci, color = elapsed_s)) +
  facet_wrap(~id, scales= "free") + 
  geom_point() 
```



```{r}
library(plantecophys)

## 2)  Load a-ci curves
a.ci <- aci_post
str(a.ci)
head(a.ci)

## 3) Use "Curve" as a unique identifier for each sample and set it as a separate vector
# Check out each curve for fit
# a.ci$Curve <- paste(a.ci$time, a.ci$population, a.ci$clone, a.ci$bed, a.ci$plant, sep="-")
# a.ci$sample <- paste(a.ci$bed, a.ci$plant)
trees <- unique(a.ci$id)
trees <- sort(trees)
trees

for (i in 1:length(trees)){
  test <- a.ci[a.ci$id == trees[i],]
  t <- fitaci(test, fitmethod="bilinear", Tcorrect=FALSE)
  plot(t, main=trees[i])
}

## 4) Store A/Ci curve as dataframe
vcmax.default <- matrix(data=NA, nrow=length(trees), ncol=6)

for (i in 1:length(trees)) {
  test <- a.ci[a.ci$id == trees[i],]
  t <- fitaci(test, fitmethod="bilinear", Tcorrect=FALSE)
  vcmax.default[i,1]<-t$pars[1,1]
  vcmax.default[i,2]<-t$pars[1,2]
  vcmax.default[i,3]<-t$pars[2,1]
  vcmax.default[i,4]<-t$pars[2,2]
  vcmax.default[i,5]<-t$pars[3,1]
  vcmax.default[i,6]<-t$pars[3,2]
}

vcmax.default <- as.data.frame(vcmax.default)
vcmax.default <- cbind(trees, vcmax.default)
colnames(vcmax.default) <- c("id", "Vcmax", "vcmax.error", "Jmax", "Jmax.error",
                     "Rd", "Rd.error")
data <- vcmax.default %>% 
  clean_names() %>% 
  separate(col = id, into = c("treatment", "rep"), sep = "_")

## 5) Summarize and graph Vcmax
vcmax.sum <- data %>% 
  group_by(treatment) %>%
  dplyr::summarise(mean.Vcmax = mean(vcmax, na.rm = TRUE),
                   median.Vcmax = median(vcmax, na.rm=TRUE),
                   sd.Vcmax = sd(vcmax, na.rm = TRUE),
                   n.Vcmax = length(vcmax[!is.na(vcmax)])) %>%
  mutate(se.Vcmax = sd.Vcmax / sqrt(n.Vcmax),
         lower.ci.Vcmax = mean.Vcmax - qt(1 - (0.05 / 2), n.Vcmax - 1) * se.Vcmax,
         upper.ci.Vcmax = mean.Vcmax + qt(1 - (0.05 / 2), n.Vcmax - 1) * se.Vcmax)

data %>% 
  ggplot(aes(y = vcmax, x = treatment, color = treatment))  +
  #geom_jitter(alpha = .3) +
  stat_summary(
    fun = mean,                         # Plot the mean as the dot
    fun.data = mean_sdl,                 # Plot error bars using standard error
    geom = "pointrange",                # Point with error bars
    size = 1,                          # Size of the dot
    position = position_dodge(width = 1)
  )  +
  theme(legend.position = "none") +
  scale_color_manual(values = c( "lightpink", "darkgreen")) +
  #facet_wrap(~date ,nrow = 1) +
  labs(y = "VCmax", 
       x = "Time") 
```


```{r}
fitaci(aci_post, 
       varnames = list(aleaf = 3, 
                       Tleaf = )
       )
```

#------

####GMIN: 

```{r}
#weights
gmin_weights <- read_csv(here::here("raw-data","SMOKE DATA", "GMIN DATA", "smoke_gmin.csv")) %>% 
  mutate(leaf = as.factor(leaf))

#leaf area
gmin_leafareas <- read_csv(here::here("raw-data","SMOKE DATA", "GMIN DATA", "leaf_areas.csv")) %>% 
  mutate(tree = as.factor(leaf)) %>% 
  select(-leaf)

#need hobo data

#combine
gmin_df <- merge(gmin_weights, gmin_leafareas, by = c("tree", "treatment"), all = T) %>% 
  filter(mass_change > 0)

gmin_df %>% 
  ggplot(aes(y = mass, x = time, color = leaf)) + 
  geom_point() +
  facet_wrap(~treatment, scales = "free")
```

```{r, eval = F}
gmin <- gmin_df %>% 
  #select(Species, ID, Leaf_Num, Time, Mass, Temp, RH, Atm_Pressure, Leaf_Area) %>% 
  #separate(col = "ID", into = c("treatment", "id"), sep = "_") %>% 
  mutate(Temp = as.numeric(Temp),
         RH = as.numeric(RH),
         Atm_Pressure = as.numeric(Atm_Pressure),
         Mass = as.numeric(Mass),
         Time = hm(Time),
         vpsat_kpa = 0.61121*exp((18.678-(Temp/234.5))*(Temp/(257.14+Temp))),
         mfvpd = (1-(RH/100))*(vpsat_kpa/Atm_Pressure)) %>% 
  group_by(Species, treatment, id, Leaf_Num) %>% 
  mutate(mass_change = lag(Mass) - Mass,
         time_change = as.duration(Time - lag(Time))) %>% 
  ungroup() %>% 
  drop_na(mass_change) %>% 
  mutate(time_change = as.numeric(time_change),
         gmin = ((mass_change)/18*1000)/(time_change)/mfvpd/(Leaf_Area*2/10000)) %>% 
  filter(gmin >= 0)

gmin_processed <- read.csv(here("data", "gmin_processed.csv"))

enca_gmin <- gmin_processed %>% 
  filter(Species == "ENCA") 

TukeyHSD(aov(gmin~treatment, data = enca_gmin))

ggplot(enca_gmin, aes(x = treatment, y = gmin))+
  geom_boxplot()+
  theme_minimal()

mala_gmin <- gmin_processed %>% 
  filter(Species == "MALA")

TukeyHSD(aov(gmin~treatment, data = mala_gmin))

ggplot(mala_gmin, aes(x = treatment, y = gmin)) +
  geom_boxplot()+
  theme_minimal()
```

#-----

#####weights


```{r}
weights_df <- read_csv(here::here("raw-data","SMOKE DATA", "Weights_Environment", "weights.csv")) %>% 
              clean_names() %>% 
  mutate(date= as.POSIXct(date, format = "%m/%d/%y"),
         date_new = lubridate::ymd(date), 
         tree_id = as.character(tree_id)
         ) %>% 
  unite(id, c("chamber_no", "tree_id")) %>% 
  group_by(date_new, id) %>% 
  pivot_wider(names_from = pre_post, 
              values_from = mass_g) %>% 
  ungroup() %>% 
  group_by(date_new, id) %>% 
  fill(pre, post, .direction = "downup") %>% 
  ungroup() %>% 
  group_by(id) %>% 
  fill(14:18, .direction = "downup")

weights_prepost_df <- weights_df %>% 
  select(date_new, id, pre, post) %>% 
  mutate(pre = as.numeric(pre), 
         post = as.numeric(post)) %>% 
  mutate(weight_change_g = pre-post) %>% 
  distinct() %>% 
  filter(weight_change_g > -5) %>% 
  separate(id, into = c("treatment", "plant_id"), sep = "smoke_") %>% 
  mutate(treatment = case_when(
    treatment %in% c("no_") ~ "nosmoke", 
    TRUE ~ "smoke"
  ))

#Just entered the wrong data, woops
weights_prepost_df %>% 
  ggplot(aes(y = weight_change_g, x = treatment, color = treatment))  +
  geom_jitter(alpha = .3) +
  stat_summary(
    fun = mean,                         # Plot the mean as the dot
    fun.data = mean_sdl,                 # Plot error bars using standard error
    geom = "pointrange",                # Point with error bars
    size = .75,                          # Size of the dot
    position = position_dodge(width = .5)
  ) +
  facet_wrap(~date_new, nrow = 1, scales = "free") +
  labs(y = "Transpirative Loss (g))", 
       x = "Time", 
       color = "Treatment")
  
```

